---
title: "Lecture 5"
author: "KM Purcell"
date: "October 13, 2015"
output: ioslides_presentation
---

## Previously
- indexing data
- filtering data frames
- sorting data frames
- producing data summaries
- contingency tables (cross tabulations)

## Objectives
- Conclude descriptive statistics
- Descriptive statistics
- Data visualization tools

## Continue descriptive statistics tools

## Descriptive statistics by groups

## Dependence vs Correlation
- tests of independence


## Chi square test

```{r, warning=FALSE}
mytable <- xtabs(~cyl+gear, data=mtcars)  #cross tabulate data
mytable  #View data
chisq.test(mytable) #evalute independence
```



## Correlation



# Descriptive Statistics




## `iris` dataset

- Today we are using a new dataset
- `iris` data is a data frame with 150 cases (rows) & 5 variables (columns) 
- Variables are Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, and Species.
- Data derived from Anderson but popularized by **RA Fisher**
- Who is Fisher?

## Visualizing with `base`

- There a number of packages in `R` that facilitate the graphical display of data.
- `base` graphics come standard with all `R` installations
- `ggplot2` and `lattice` are two additional but standard packages

## Getting to know `iris`
```{r}
# New data set 
str(iris)

# Names
names(iris)
```

## Getting to know `iris`

```{r}
summary(iris)
```




