---
title: "Lecture 6 - Exploratory Data Analysis"
date: "October 17, 2015"
output: 
  revealjs::revealjs_presentation:
     theme: serif
     center: true
     self_contained: true
     highlight: default
---

## Lecture 6

- Conclude an introduction to EDA
- Review research proposals
- Work on R assignment


# Exploratory Data Analysis

## Boxplots
- Graphical depictions of groups of numerical data through their quartiles. 
- Box plots may have lines extending indicating variability outside the upper and lower quartiles
- Hence the terms box-and-whisker plot and box-and-whisker diagram.

##
```{r}
# Boxplot of MPG by Car Cylinders 
boxplot(mpg~cyl,data=mtcars, main="Car Milage Data", 
  	xlab="Number of Cylinders", ylab="Miles Per Gallon")
```


##
```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=6}
# Notched Boxplot of Tooth Growth Against 2 Crossed Factors
# boxes colored for ease of interpretation 
boxplot(len~supp*dose, data=ToothGrowth, notch=TRUE, 
  col=(c("gold","darkgreen")),
  main="Tooth Growth", xlab="Suppliment and Dose")
```

## Violin Plots
- Blah Blah
- Blah Blah

##
```{r, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# Violin Plots
library(vioplot); 
x1 <- mtcars$mpg[mtcars$cyl==4];x2 <- mtcars$mpg[mtcars$cyl==6];x3 <- mtcars$mpg[mtcars$cyl==8]
vioplot(x1, x2, x3, names=c("4 cyl", "6 cyl", "8 cyl"), 
   col="gold")
title("Violin Plots of Miles Per Gallon")
```



## QQ Plots
is a probability plot, which is a graphical method for comparing two probability distributions by plotting their quantiles against each other. 

##
```{r}
# Q-Q Plot for variable MPG 
qqnorm(mtcars$mpg)
qqline(mtcars$mpg)
```

##
```{r, warning=FALSE, message=FALSE}
library(car); library(multcomp)
qqPlot(lm(response~ trt, data=cholesterol), 
       simulate=TRUE, main="Q-Q Plot", labels=FALSE)

```


## Dot plots
- a representation of a distribution with groups of data points plotted on a simple scale. 
- Dot plots are used for continuous, quantitative, univariate data. 





##
```{r, fig.width=6, fig.height=5.5}
# Simple Dotplot
dotchart(mtcars$mpg,labels=row.names(mtcars),cex=.7,
  	main="Gas Milage for Car Models", 
   xlab="Miles Per Gallon")
```

##
```{r, echo=FALSE}
# Kabacoff 2015, p. 134
x <- mtcars[order(mtcars$mpg),]
x$cyl <- factor(x$cyl)

x$color[x$cyl==4] <- "red"
x$color[x$cyl==6] <- "blue"
x$color[x$cyl==8] <- "darkgreen"

dotchart(x$mpg,
         labels = row.names(x),
         cex = .7,
         groups = x$cyl,
         gcolor = "black",
         color = x$color,
         pch=19,
         main = "Gass Mileage for Car Models\ngrouped by cylinder",
         xlab = "Miles per Gallon")
```

- sorted by `mpg`, grouped & colored by `cyl`

##
